### 2021-02-24 manuscript/manuscript-speed-accuracy.tex
Updated README.md, manuscript and figures.

\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{amsmath}
\graphicspath{{../figures/}}
\setlength{\abovecaptionskip}{15pt plus 3pt minus 2pt}

@@

\Authors{Paul P. Gardner\textsuperscript{1,2}*, James M. Paterson\textsuperscript{?}, Fatemeh Ashari-Ghomi\textsuperscript{?}, Sinan U. Umu\textsuperscript{?}, Stephanie McGimpsey\textsuperscript{3,4}, Aleksandra Pawlik\textsuperscript{5},
Alex Gavryushkin\textsuperscript{6}, Michael A Black\textsuperscript{1}
} % Authors

\affiliation{\textsuperscript{1}\textit{Department of Biochemistry, University of Otago, Dunedin, New Zealand.}} % Author affiliation
\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
\affiliation{\textsuperscript{3}\textit{McConomy School of Dance, Derry, Ireland.}}
\affiliation{\textsuperscript{4}\textit{Research for Good, Limavady, Ireland.}}
\affiliation{\textsuperscript{5}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
\affiliation{\textsuperscript{6}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation
\affiliation{*\textbf{Corresponding author}: paul.gardner@otago.ac.nz} % Corresponding author

@@

%\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
\Abstract{ Computational biology has provided widely used and powerful
  software tools for testing and making inferences about biological
  data. In the face of rapidly increasing volumes of data, heuristic
  methods that trade software speed for mathematical completeness must
  be employed.  We are interested in testing for trade-offs between
  speed and accuracy and for other factors are indicative of accurate
  software.

We have extracted accuracy and speed ranks from independent benchmarks of
computational biology software of different software tools, and evaluate the
factors that are likely to influence accuracy e.g. speed, author
reputation, journal impact, recency or developer efforts.

We found that software speed, author reputation, journal impact, the
number of citations and age are all unreliable predictors of software
accuracy. This is unfortunate because citations, author and journal
reputation are frequently used reasons for selecting software
tools. Our results do show that accurate bioinformatic software tools
are generally the product many gradual improvements, often from
multiple developers, based upon github records. In addition, we find
that bioinformatics has an excess of slow and inaccurate software
tools across many sub-disciplines. Meanwhile, there are few tools that
are middle-of-road in terms of accuracy and speed trade-offs. We
hypothesise that a form of publication-bias unduly influences the
publication and development of bioinformatic software tools. In other
words, at present software that is not highly ranked on speed and not
highly ranked on accuracy is difficult to publish due to author,
editor and reviewer practices. This leaves an unfortunate gap in the
literature upon which future software refinements cannot be
constructed.  }

@@

\section*{Background}
Computational biology software is widely used and has produced some of
the most cited publications in the entire scientific corpus
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. This
highly-cited software includes implementations of methods for sequence
alignment and homology inference
\cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
phylogenetic analysis
\cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei},
statistical analysis of survival patterns in biomedicine
\cite{Kaplan1958-ju,Cox1972-nu}, biomolecular structure analysis
\cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj},
visualization and data collection
\cite{Kraulis1991-lt,Berman2000-to}. However, the popularity of a
software tool does not necessarily mean that it is accurate or
computationally efficient, instead usability, ease of installation,
operating system and other factors may play a greater role in a software
tool's popularity. There have been several notable incidences where
inaccurate software has caused considerable harm
\cite{leveson1993investigation}.
%Fergusson covid predictions?

Progress in the biological sciences is increasingly limited by the
ability to analyse increasing volumes of data, therefore the
dependence of biologists on software is also increasing
\cite{Marx2013-zi}. There is an increasing use of technological
solutions for automating biological data generation
(e.g. next-generation sequencing, mass-spectroscopy, cell-tracking and
species tracking), therefore the biological sciences have become
increasingly dependent upon computational software for processing
large quantities of data \cite{Marx2013-zi}. As a consequence, the
computational efficiency of data processing and analysis software is
of great importance to decrease the energy and time costs of research
\cite{Gombiner2011-md}. Furthermore, even small error rates can have a
major impact on the number of false inferences as datasets become
larger \cite{Storey2003-cv}.

@@

For common computational biology tasks, a proliferation of
software-based solutions often exists
\cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut,Wikipedia_contributors2015-vj,Wikipedia_contributors2015-hr}. While
this is a good problem to have, and points to a
diversity of options from which practical solutions can be selected,
many possible options creates a dilemma for users. In the absence of
any recent gold-standard benchmarks, how should scientific software be
selected? In the following we presume that ``biological
accuracy'' is the most desirable feature for a software tool. Biological
accuracy is the degree to which predictions or measurements reflect
the biological truths based on expert-derived curated datasets. In some
fields biological accuracy is very difficult to ascertain, for
example, in phylogenetics it is nearly impossible to know the
true ancestral relationships between organisms. In situations like this,
researchers may use simulated or high-confidence datasets.

A number of possible predictors of software quality are used by the
community of computational biology software users. Some accessible,
quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work or be an update to
an existing software. Therefore, these could be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact, “large user base”, “word-of-mouth”,
“wide-adoption”, "personal recommendation," and "recommendation from a
close colleague," are frequent responses to surveys of “how do
scientists select software?”
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
  impact:} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/Group
  reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors will produce better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromise
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph}; however, the implementation is
likely to have more of an impact (e.g. brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo}).

Other factors that are less quantifiable that influence whether a
software tool is selected include: whether the documentation is good,
user-friendly, word-of-mouth and ``used in a similar analysis''
\cite{Loman2015-bw}.
%This information is not as readily
%quantifiable as the above measures. However,
However, citation metrics may be a
useful proxy for the above.
%This may also explain the reason why
%some software continues to be used, in spite of poor relative
%performance \cite{Wadi2016-dj}.
%We note that none of these factors
%are directly linked to the top ranked features of successful software
%projects, as based upon literature surveys \cite{nasir2011critical}.

With the wide adoption of Github ($\approx 47\%$ of the
\textbf{{\color{red}\numTools}} tools included in this study could be
linked to a Github repository), quantifiable data on software
development such as the number of contributors to code, number of code
changes and versions is also available
\cite{ray2014large,Dozmorov:2018,mangul2018comprehensive}.

In the following study, we explore factors that may be indicative of
software accuracy. This, in our opinion, should be one of the prime
reasons for selecting a software tool. We have mined the large and
freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks
of computational biology software, and manually extracted accuracy and
speed rankings for \textbf{{\color{red}\numTools~}} unique software tools. For
each tool, we have collected measures that may be predictive
of accuracy, and may be subjectively employed by the researcher
community as a proxy for software quality. These include relative
speed, relative age, the productivity and impact of the corresponding
authors, journal impact and the number of citations.

\section*{Results}
We have collected relative accuracy and speed ranks for
\textbf{{\color{red}\numTools~}} distinct software tools. This software has
been developed for solving a broad cross-section computational biology
tasks.
%% These include software for homology search
%% \cite{Freyhult2007-et}, genome sequence analysis (e.g. read mapping or
%% sequence assembly)
%% \cite{Junemann2014-mb,Tran2014-pe,Zhang2011-nd,Abbas2014-gu,Bao2011-lv,Caboche2014-lj,Kleftogiannis2013-wi,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Yang2013-aj,Holtgrewe2011-fd,Rackham2015-ag,Huang2015-wu},
%% multiple sequence alignment
%% \cite{Thompson2011-rf,Nuin2006-nk,Pais2014-sr,Pervez2014-zp,Liu2010-rp},
%% cell tracking \cite{Maska2014-ak}, transcriptome analysis
%% \cite{Li2012-wr,Lu2013-fs,Liu2014-kz,Kumar2016-xz}, RNA interaction
%% prediction \cite{Pain2015-gr}, protein interactions
%% \cite{Tikk2010-qd}, protein structure prediction
%% \cite{Kolodny2005-ry,Wallner2005-qi}, epistasis \cite{Shang2011-vy},
%% metagenomic analysis \cite{Lindgreen2016-tt,Bazinet2012-wf},
%% repetitive sequence prediction \cite{Saha2008-kd}, proteomics
%% \cite{Lange2008-pt,Yang2009-oc} and phylogenetics
%% \cite{Liu2011-pz,Yang2011-dv,Oscamou2008-md,Bayzid2013-hc,Liu2009-lx}.
Each
software tool was benchmarked in at least one of
\textbf{{\color{red}\numBenchmarkPubs~}} publications that satisfy the Boulesteix
criteria \cite{Boulesteix2013-vb}. In brief, the Boulesteix criteria
are: 1. the main focus of the article is a benchmark. 2. the authors
are reasonably neutral. 3. the test data and evaluation criteria are
sensible.

For each of the publications describing these tools, we have (where
possible) the H5-index published by
\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{Google
  Scholar Metrics}. We have collected the maximum H-index and
corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
authors for each tool, and the number of times the publication(s)
associated with a tool has been cited using Google Scholar (data
collected over a 6 month period in late 2020). Note that the citation metrics
are not static and will change over time.

We have computed the Spearman’s correlation coefficient for each
pairwise combination of the mean normalised accuracy and speed ranks,
the year published, mean relative age (compared to software in the
same benchmarks), journal H5 metrics, the total number of citations,
the relative number of citations (compared to software in the same
benchmarks) and the maximum H- and corresponding M-indices for the corresponding
authors. Where possible we also extract the version number, the number
of commits and number of contributors from Github repositories.  The
results are presented in Figure~\ref{fig:allfactors}A. We find
significant associations between most of the citation-based metrics
(journal H5, citations, relative citations, H-index and
M-index). There is also a negative correlation between the year of
publication, the relative age and many of the citation-based metrics.

\begin{figure*}[h]
\includegraphics[width=\textwidth]{figure1.pdf}
\caption{\textbf{A.} A heatmap indicating the relationships between
  proposed predictors of software quality. Spearman’s rho is used to
  infer correlations between metrics such as the H- and M-indices of
  corresponding authors, number of citations, journal impact factors
  and H5 indices, the year and relative age of software and the mean
  relative rankings of software speed and accuracy. Red colours
  indicate a positive correlation, blue colours indicate a negative
  correlation. Correlations with a P-value less than 0.05 are
  indicated with a ‘X’. The dashed rectangular area is illustrated in
  more detail in \textbf{B}, the relationship between speed and
  accuracy is shown in more detail in \textbf{Figure 2}.
  %The
  %dendrogram was computed using the default ‘heatmap.2 function in R,
  %which computes a Euclidean distance matrix and a complete-linkage
  %hierarchical clustering.
  \textbf{B.} A barplot illustrating the correlation, as measured by
  Spearman’s rho, between normalised accuracy ranks and software tool
  features that may be predictive of accuracy. In order to give an
  appreciation of the difference between the observed effect-sizes and
  significant effect sizes, we generated 1000 permuted accuracy ranks
  for each benchmark and recorded Spearman’s rho for the significant
  correlations (P<0.05). These values are marked with “x”s in the
  barplot. Significant correlations are marked with red asterisks.}
\label{fig:allfactors}
\end{figure*}

Data on the
number of updates to software tools from Github such as the number of versions,
commits and contributors \textbf{{\color{red}was significantly correlated with 
accuracy (respective Spearman’s rhos = 0.13, 0.22, 0.21, respective P-values = 0.0027, 0.00079, 0.0013). 
These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A).
}}
We also found that author reputation metrics, journal impacts and the age
of tools were generally \textbf{not} correlated with either tool
accuracy or speed (see Figure~\ref{fig:allfactors}). The strongest
association was between accuracy and \textbf{{\color{red}the relative
    number of citations}} (Spearman’s rho = \textbf{{\color{red}0.092,
    P-value = 0.047}}). But the effect size is only just over the
significance threshold and may be explained by multiple-testing.

@@

In order to gain a deeper understanding of the distribution of
available bioinformatic software tools on a speed versus accuracy
landscape, we ran a permutation test. The ranks extracted
from each benchmark were randomly permuted, generating 1,000
randomized speed and accuracy ranks. In the cells of a $10\times10$
grid spanning the normalised speed and accuracy ranks we computed a
Z-score for the observed number of tools in a cell, compared to the
expected distributions generated by 1,000 randomized ranks. The results
of which are shown in Figure~\ref{fig:speedaccuracy}. We identified \textbf{{\color{red}18}}
bins where there was a significant excess or dearth of tools. For
example, there was an excess of “slow and inaccurate” software (\textbf{{\color{red}Z=2.50,
P-value=0.0063}}). We find that the amount of software classed as “fast
and accurate” and “fast and inaccurate” are at approximately the
expected proportions based upon the permutation test. The number of
significant results is in excess, and is not due to multiple testing, as
the probability of finding \textbf{{\color{red}18}} of 100 tests significant by chance is
low (P-value = \textbf{{\color{red}$2.2\times 10^{-6}$}}, exact binomial test).

@@

\begin{figure*}
%\includegraphics[scale=0.55]{figure2.pdf}
\includegraphics[width=\textwidth]{figure2.pdf}
\caption{A heatmap indicating the relative paucity or abundance of
  software in the range of possible accuracy and speed rankings. Red
  colours indicate an abundance of software tools in an accuracy and
  speed category, while blue colours indicate scarcity of software in
  an accuracy and speed category. The abundance is quantified using a
  Z-score computation for each bin, this is derived from 10,000 random
  permutations of the speed and accuracy ranks from each
  benchmark. Mean normalised ranks of accuracy and speed have been
  binned into 100 classes (a $10\times10$ grid) that range from
  comparatively slow/inaccurate to comparatively
  fast/accurate. Z-scores with a P-value less than 0.05 are indicated
  with a ‘X’.}
\label{fig:speedaccuracy}
\end{figure*}

@@

We have gathered data on the relative speeds and accuracies of
\textbf{{\color{red}\numTools~}} bioinformatic tools from
\textbf{{\color{red}\numBenchmarkPubs}} benchmarks published between
2005 and 2020. {\bf Our results suggest there are major benefits for
  long term support for software \cite{siepel2019challenges}.} The
finding of a strong relationship between the number of commits to
github (i.e. software updates) and accuracy, highlights the benefits of
long-term development.

Our study finds little evidence to support that impact-based metrics
are poorly related to software quality, this is unfortunate as these
are frequently cited reasons for selecting software tools
\cite{Loman2015-bw}. This implies that the recorded high citation
rates for bioinformatic software
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
reflection of user-friendliness and possibly the Matthew Effect
\cite{Lariviere2010-kx,Merton1968-cb} than accuracy.

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are overrepresented at both
high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
In addition, we find an under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1959publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences.  We also found that slow and inaccurate
software is generally published earlier than fast and accurate tools
\textbf{{\color{red}(P=0.007, one-tailed Wilcoxon test)}}, therefore
comparisons were not required to publish the slow and
inaccurate tools.

@@

PPG is supported by a Rutherford Discovery Fellowship,
administered by the Royal Society of New Zealand. 
{\bf AND ETC....}

>>>>

\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{todonotes}
\usepackage{maori}
\graphicspath{{../figures/}}
\setlength{\abovecaptionskip}{15pt plus 3pt minus 2pt}

@@

\Authors{Paul P. Gardner\textsuperscript{1,2}*, James M. Paterson\textsuperscript{3}, Stephanie McGimpsey\textsuperscript{4}, Fatemeh Ashari-Ghomi\textsuperscript{5}, Sinan U. Umu\textsuperscript{6}, Aleksandra Pawlik\textsuperscript{7},
Alex Gavryushkin\textsuperscript{8}, Michael A Black\textsuperscript{1}
} % Authors

\affiliation{\textsuperscript{1}\textit{Department of Biochemistry, University of Otago, Dunedin, New Zealand.}} % Author affiliation
\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
\affiliation{\textsuperscript{3}\textit{Department of Civil and Natural Resources Engineering, University of Canterbury, Christchurch, New Zealand.}}
\affiliation{\textsuperscript{4}\textit{Parasites and Microbes, Wellcome Sanger Institute, Wellcome Genome Campus, Hinxton,
8 Cambridgeshire, CB10 1RQ, UK.}}
\affiliation{\textsuperscript{5}\textit{Research Group for Genomic Epidemiology, National Food Institute, Technical University of Denmark, Kongens Lyngby, Denmark.}}
\affiliation{\textsuperscript{6}\textit{Department of Research, Cancer Registry of Norway, Oslo, Norway.}} % Author affiliation
\affiliation{\textsuperscript{7}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
\affiliation{\textsuperscript{8}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation

\affiliation{*\textbf{Corresponding author}: paul.gardner@otago.ac.nz} % Corresponding author

@@

%\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
\Abstract{ Computational biology has provided widely used and powerful
  software tools for testing and making inferences about biological
  data. In the face of rapidly increasing volumes of data, heuristic
  methods that trade software speed for accuracy must
  be employed.  We are interested in testing for trade-offs between
  speed and accuracy and for other factors that are indicative of accurate
  software.

We have extracted accuracy and speed ranks from independent benchmarks of
computational biology software of different software tools, and evaluated the
factors that are likely to influence accuracy e.g. speed, author
reputation, journal impact, recency or developer efforts.

We found that software speed, author reputation, journal impact, the
number of citations and age are all unreliable predictors of software
accuracy. This is unfortunate because citations, author and journal
reputation are frequently used reasons for selecting software
tools. Our results do show that accurate bioinformatic software tools
are generally the product of many gradual improvements, often from
multiple developers, based upon matching GitHub records with benchmark results. 
In addition, we find
that the field of bioinformatics has an excess of slow and inaccurate software
tools, this is consistent across many sub-disciplines. Meanwhile, there are few tools that
are middle-of-road in terms of accuracy and speed trade-offs. We
hypothesise that a form of publication-bias unduly influences the
publication and development of bioinformatic software. In other
words, software that is intermediate in terms of both speed and accuracy may be difficult to publish due to author,
editor and reviewer practices. This leaves an unfortunate gap in the
literature as the ideal tools may fall in the gap e.g. high accuracy is generally not useful if years of CPU time is required, 
while high speed is not useful if the results are very inaccurate.  }

@@

\section*{Background}
Computational biology software is widely used and has produced some of
the most cited publications in the entire scientific corpus
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. This
highly-cited software includes implementations of methods for sequence
alignment and homology inference
\cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
phylogenetic analysis
\cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei},
statistical analysis of survival patterns in biomedicine
\cite{Kaplan1958-ju,Cox1972-nu}, biomolecular structure analysis
\cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj},
visualization and data collection
\cite{Kraulis1991-lt,Berman2000-to}. However, the popularity of a
software tool does not necessarily mean that it is accurate or
computationally efficient, instead usability, ease of installation,
operating system and other factors may play a greater role in a software
tool's popularity. Indeed, there have been several notable incidences where
inaccurate software has caused considerable harm
\cite{leveson1993investigation,cummings2020regulating,Herkert:2020}.
%Fergusson covid predictions?

Progress in the biological sciences is increasingly limited by the
ability to analyse large volumes of data, therefore the
dependence of biologists on software is also increasing
\cite{Marx2013-zi}. There is an increasing reliance on technological
solutions for automating biological data generation
(e.g. next-generation sequencing, mass-spectroscopy, cell-tracking and
species tracking), therefore the biological sciences have become
increasingly dependent upon computational software for processing
large quantities of data \cite{Marx2013-zi}. As a consequence, the
computational efficiency of data processing and analysis software is
of great importance to decrease the energy, climate impact, and time costs of research
\cite{Gombiner2011-md}. Furthermore, even small error rates can have
major impacts on the number of false inferences as datasets become
larger \cite{Storey2003-cv}.

@@

For common computational biology tasks, a proliferation of
software-based solutions often exists
\cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut,Wikipedia_contributors2015-vj,Wikipedia_contributors2015-hr}. While
this is a good problem to have, and points to a
diversity of options from which practical solutions can be selected,
having many possible options creates a dilemma for users. In the absence of
any recent gold-standard benchmarks, how should scientific software be
selected? In the following we presume that ``biological
accuracy'' is the most desirable feature for a software tool. Biological
accuracy is the degree to which predictions or measurements reflect
the biological truths based on expert-derived curated datasets (see Methods for the  mathematical definition used here). In some
fields biological accuracy is very difficult to ascertain, for
example, in phylogenetics it is nearly impossible in most cases to know the
true ancestral relationships between organisms. In situations like this,
researchers may use simulated or high-confidence datasets.

A number of possible predictors of software quality are used by the
community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. Some accessible,
quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work, or be an update to
an existing software. Therefore these may be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact, “large user base”, “word-of-mouth”,
“wide-adoption”, "personal recommendation," and "recommendation from a
close colleague," are frequent responses to surveys of “how do
scientists select software?”
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
  impact:} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/Group
  reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors will produce better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromises
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph};
however, the implementation generally has more of an impact (e.g. brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo}).
%\todo{This paper
%https://science.sciencemag.org/content/368/6495/eaam9744.full
%shows a thousands-fold increase in speed when parallelisation
%is implemented with things like cache, architecture, etc.
%mind VS naive parallelisation as in Python, R, etc.}

Other factors that are less quantifiable that influence whether a
software tool is selected include: whether the documentation is good,
user-friendly, word-of-mouth and ``used in a similar analysis''
\cite{Loman2015-bw}. However, citation metrics may be a
useful proxy for the above.
%This may also explain the reason why
%some software continues to be used, in spite of poor relative
%performance \cite{Wadi2016-dj}.
%We note that none of these factors
%are directly linked to the top ranked features of successful software
%projects, as based upon literature surveys \cite{nasir2011critical}.

With the wide adoption of GitHub ($\approx 47\%$ of the
\textbf{{\color{red}\numTools}} tools included in this study could be
linked to a GitHub repository), quantifiable data on software
development such as the number of contributors to code, number of code
changes and versions is also available
\cite{ray2014large,Dozmorov:2018,mangul2018comprehensive}.

In the following study, we explore factors that may be indicative of
software accuracy. This, in our opinion, should be one of the prime
reasons for selecting a software tool. We have mined the large and
freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks
of computational biology software, and manually extracted accuracy and
speed rankings for \textbf{{\color{red}\numTools~}} unique software tools. For
each tool, we have collected measures that may be predictive
of accuracy, and may be subjectively employed by the research
community as a proxy for software quality. These include relative
speed, relative age, the productivity and impact of the corresponding
authors, journal impact and the number of citations.

\section*{Results}
We have collected relative accuracy and speed ranks for
\textbf{{\color{red}\numTools~}} distinct software tools. This software has
been developed for solving broad cross-section computational biology
tasks.
Each software tool was benchmarked in at least one of
\textbf{{\color{red}\numBenchmarkPubs~}} publications that satisfy the Boulesteix
criteria \cite{Boulesteix2013-vb}. In brief, the Boulesteix criteria
are: 1. the main focus of the article is a benchmark. 2. the authors
are reasonably neutral. 3. the test data and evaluation criteria are
sensible.

For each of the publications describing these tools, we have (where
possible) collected the journal's H5-index (\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{Google
  Scholar Metrics}), the maximum H-index and
corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
authors for each tool, and the number of times the publication(s)
associated with a tool has been cited using Google Scholar (data
collected over a 6 month period in late 2020). Note that citation metrics
are not static and will change over time. In addition, where possible we also extract the version number, the number
of commits and number of contributors from GitHub repositories.

We have computed the Spearman’s correlation coefficient for each
pairwise combination of the mean normalised accuracy and speed ranks,
with the year published, mean relative age (compared to software in the
same benchmarks), journal H5 metrics, the total number of citations,
the relative number of citations (compared to software in the same
benchmarks) and the maximum H- and corresponding M-indices for the corresponding
authors, version number, and numbers of commits and contributors.
The results are presented in Figure~\ref{fig:allfactors}A. We find
significant associations between most of the citation-based metrics
(journal H5, citations, relative citations, H-index and
M-index). There is also a negative correlation between the year of
publication, the relative age and many of the citation-based metrics.

\begin{figure*}[ht]
\includegraphics[width=\textwidth]{figure1.pdf}
\caption{\textbf{A.} A heatmap indicating the relationships between
  proposed predictors of software quality. Spearman’s rho is used to
  infer correlations between metrics such as the H- and M-indices of
  corresponding authors, number of citations, journal impact factors
  and H5 indices, the year and relative age of software and the mean
  relative rankings of software speed and accuracy. Red colours
  indicate a positive correlation, blue colours indicate a negative
  correlation. Correlations with an uncorrected P-value less than 0.05 are
  indicated with a `$\vartimes$'. The dashed rectangular area is illustrated in
  more detail in \textbf{B}, the relationship between speed and
  accuracy is shown in more detail in \textbf{Figure 2}.
  %The
  %dendrogram was computed using the default ‘heatmap.2 function in R,
  %which computes a Euclidean distance matrix and a complete-linkage
  %hierarchical clustering.
  \textbf{B.} A barplot illustrating the correlation, as measured by
  Spearman’s rho, between normalised accuracy ranks and software tool
  features that may be predictive of accuracy. In order to give an
  appreciation of the difference between the observed effect-sizes and
  significant effect sizes, we generated 1000 permuted accuracy ranks
  for each benchmark and recorded Spearman’s rho for the significant
  correlations with an uncorrected P-value less than 0.05.
  These values are marked with blues `$\vartimes$'s in the
  barplot. Significant correlations are marked with red asterisks.}
\label{fig:allfactors}
\end{figure*}

Data on the
number of updates to software tools from GitHub such as the number of versions,
commits and contributors \textbf{{\color{red}was significantly correlated with 
accuracy (respective Spearman’s rhos = 0.13, 0.22, 0.21, respective uncorrected P-values = 0.0027, 0.00079, 0.0013). 
These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A).
}}
We also found that author reputation metrics, journal impacts and the age
of tools were generally \textbf{not} correlated with either tool
accuracy or speed (see Figure~\ref{fig:allfactors}). The strongest
association was between accuracy and \textbf{{\color{red}the relative
    number of citations}} (Spearman’s rho = \textbf{{\color{red}0.092,
    P-value = 0.047}}). But the effect size is only just over the
significance threshold and may be explained by multiple-testing.

@@

In order to gain a deeper understanding of the distribution of
available bioinformatic software tools on a speed versus accuracy
landscape, we ran a permutation test. The ranks extracted
from each benchmark were randomly permuted, generating 1,000
randomized speed and accuracy ranks. In the cells of a $10\times10$
grid spanning the normalised speed and accuracy ranks we computed a
Z-score for the observed number of tools in a cell, compared to the
expected distributions generated by 1,000 randomized ranks. The results
of this are shown in Figure~\ref{fig:speedaccuracy}. We identified \textbf{{\color{red}18}}
bins where there was a significant excess or dearth of tools. For
example, there was an excess of “slow and inaccurate” software (\textbf{{\color{red}Z=2.50,
P-value=0.0063}}). We find that the amount of software classed as “fast
and accurate”, “fast and inaccurate” and "slow and accurate" are at approximately the
expected proportions based upon the permutation test. 
While there are very few tools placed in the middle of the ranks on accuracy and speed.
The number of
significant results is in excess, and is not due to multiple testing, as
the probability of finding \textbf{{\color{red}18}} of 100 tests significant by chance is
low (P-value = \textbf{{\color{red}$2.2\times 10^{-6}$}}, exact binomial test).

@@

\begin{figure*}[ht]
%\includegraphics[scale=0.55]{figure2.pdf}
\centering
\includegraphics[width=0.75\textwidth]{figure2.pdf}
\caption{A heatmap indicating the relative paucity or abundance of
  software in the range of possible accuracy and speed rankings. Red
  colours indicate an abundance of software tools in an accuracy and
  speed category, while blue colours indicate scarcity of software in
  an accuracy and speed category. The abundance is quantified using a
  Z-score computation for each bin, this is derived from 10,000 random
  permutations of the speed and accuracy ranks from each
  benchmark. Mean normalised ranks of accuracy and speed have been
  binned into 100 classes (a $10\times10$ grid) that range from
  comparatively slow/inaccurate to comparatively
  fast/accurate. Z-scores with a P-value less than 0.05 are indicated
  with a ‘$\vartimes$’.}
\label{fig:speedaccuracy}
\end{figure*}

@@

We have gathered data on the relative speeds and accuracies of
\textbf{{\color{red}\numTools~}} bioinformatic tools from
\textbf{{\color{red}\numBenchmarkPubs}} benchmarks published between
2005 and 2020. {\bf Our results suggest there are major benefits for
  long term support for software \cite{siepel2019challenges}.} The
finding of a strong relationship between the number of commits to
GitHub (i.e. software updates) and accuracy, highlights the benefits of
long-term development.

Our study finds little evidence to support that impact-based metrics
have any relationship with software quality, this is unfortunate as these
are frequently cited reasons for selecting software tools
\cite{Loman2015-bw}. This implies that the high citation
rates for bioinformatic software
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
reflection of other factors such as user-friendliness or the Matthew Effect
\cite{Lariviere2010-kx,Merton1968-cb} other than accuracy.

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are overrepresented at both
high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
In addition, we find an under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1995publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences.  We also found that slow and inaccurate
software is generally published earlier than fast and accurate tools
\textbf{{\color{red}(P=0.007, one-tailed Wilcoxon test)}}, therefore
comparisons were not required to publish the slow and
inaccurate tools.

@@

PPG is supported by a Rutherford Discovery Fellowship,
administered by the Royal Society Te Ap\=arangi, 
PPG, AG and MB acknowledge support from a Data Science Programmes grant
(UOAX1932).

{\bf AND ETC....}
